### 2020-06-10 Output/manuscript/HI_Model_ms.Rmd
commit files

---
title: "Ecosystem carbon balance in the Hawaiian Islands under different scenarios of future climate and land use change"
bibliography: HI_Model.bib
csl: environmental-research-letters.csl
fontsize: 12pt
geometry: margin=1in
header-includes:
   - \usepackage{lineno}
   - \linenumbers
   - \usepackage{setspace}
   - \doublespacing
latex_engine: pdflatex  
output: pdf_document
---

```{r setup, include=FALSE}
options(tinytex.verbose = TRUE)
knitr::opts_chunk$set(echo = FALSE)
```
Paul C. Selmants^1,6^, Benjamin M. Sleeter^2^,  Jinxun Liu^1^, Tamara S. Wilson^1^,  
Parker C. Trauernicht^3^, Abby G. Frazier^4^, Gregory P. Asner^5^

**Affiliations:**  
^1^U.S. Geological Survey, Moffett Field, CA, USA   
^2^U.S. Geological Survey, Seattle, WA, USA  
^3^University of Hawaii at Manoa, Honolulu, HI, USA  
^4^The East-West Center, Honolulu, HI, USA  
^5^Arizona State University, Tempe, AZ, USA  
^6^Author to whom correspondence should be addressed

**Email:** pselmants@usgs.gov

**Running title:** Hawaii carbon balance  
**Keywords:** land use, climate change, carbon balance, Hawaii, scenarios, disturbance, ecosystem model

**Date:** `r format(Sys.time(), '%B %d, %Y')`

\pagebreak

## Abstract

The State of Hawaii recently passed legislation setting a goal to be carbon neutral by 2045. Meeting this goal will partly depend on carbon sequestration by terrestrial ecosystems, yet the future direction and magnitude of the land carbon sink in the Hawaiian Islands is highly uncertain. We used simulation modeling to assess how projected future changes in climate and land use will influence ecosystem carbon balance in the Hawaiian Islands under four unique scenarios over a 90-year timespan. Net ecosystem carbon balance declined under all four scenarios. Moving from a high to a low radiative forcing scenario reduced net ecosystem carbon loss by ~21%, and net carbon losses were reduced by a total of ~55% under the combined scenario of low radiative forcing and low rates of land‐use change. A sensitivity test of the CO~2~ fertilization effect on plant productivity revealed it to be a major source of uncertainty in projections of ecosystem carbon balance. Reconciling this uncertainty in how net photosynthesis will respond to rising atmospheric CO~2~ will be essential to better constrainment of models used to evaluate the effectiveness of ecosystem‐based climate mitigation strategies.

## Introduction

## Methods

### *Study area*

The study area encompasses the terrestrial portion of the seven main Hawaiian Islands, a total land area of 16,464 km^2^ (Figure 1).

```{r fig.align = 'center', out.width = "80%", fig.cap = "Land cover map of the seven main Hawaiian Islands, adapted from Jacobi et al. (2017). Agriculture in this map combines herbaceous and woody crops, but these two crop types are treated as separate State Classes in the simulation model."}
knitr::include_graphics("./fig_images/fig1_HI_LC.png")
```

### *States and transitions*

### *Carbon stocks and flows*

### *Initial conditions*

### *Scenario simulations*

## Results

## Discussion

## Conclusion

## Acknowledgements